---
title: "Math4Fish Entregable TODO"

author:
- Jamie Lentin^[lentinj@shuttlethread.com]
- Alfonso Perez^[alfonso.perez@ieo.csic.es]
- María José Zúñiga^[maria.zuniga@ieo.csic.es]
- María Angeles Gamaza
- Andrea Jiménez
- Margarita María Rincón
output:
  # word_document:
  #   toc: no
  #   toc_depth: '6'
  pdf_document:
    keep_tex: yes
    toc: no
    toc_depth: '4'
    number_sections: yes
header-includes:
- \usepackage{fancyhdr}
- \usepackage{array}
- \usepackage{booktabs}
- \usepackage{threeparttablex}
- \pagestyle{fancyplain}
- \usepackage{caption}
- \usepackage{listings}% http://ctan.org/pkg/listings
- \captionsetup[table]{name=Tabla,labelsep=period}
- \captionsetup[figure]{name=Figura,labelsep=period}
- \captionsetup{justification=justified,format=plain,font=small,margin=1pt}
- \fancyhead{}
- \addtolength{\headheight}{1.5cm}
- \rhead{\includegraphics[width=3.5cm]{input/LOGO_2SN_recortado.png}}
- \usepackage{float}
- \fancyfoot[R]{\thepage}
- \fancyfoot[C]{\includegraphics[height=3cm]{input/Banners_Convenio.jpeg}}
- \fancyfoot[L]{}
 #\chead{\includegraphics[height=2.0cm]{logos_imagenes/Header_logos.jpg}}
- \pagenumbering{gobble}
- \pagenumbering{arabic}
- \usepackage{fvextra}
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
always_allow_html: true
lang: es-ES
geometry: "left=3cm,right=3cm,top=2cm,bottom=4.5cm"
urlcolor: blue
csl: ices-journal-of-marine-science.csl
bibliography: referencias.bib
link-citations: yes
# editor_options:
#   markdown:
#     wrap: 72
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=T, warning=FALSE, message=FALSE,comment="",collapse=TRUE,fig.align="center",fig.pos="h!",eval=F,tidy=TRUE, tidy.opts=list(arrow=TRUE, indent=2))
```


```{r results='hide',include=FALSE,eval=T}
## Librerías requeridas
 paquetes <- c("stringr", "tidyverse", "kableExtra","ggplot2","ggthemes",
               "patchwork","dplyr","reshape","here","r4ss","zoo")
 lapply(paquetes, require, character.only = TRUE)
```


<!-- \mbox{} con esto inserto una línea en blanco -->

\mbox{}

\mbox{}

\mbox{}

Instituto Español de Oceanografía, Consejo Superior de Investigaciones
Científicas (IEO-CSIC)

```{=tex}
\renewcommand{\tablename}{Tabla}
\newpage
```
\mbox{} \setcounter{tocdepth}{5} \tableofcontents \newpage

# Math4Fish User guide

## Overview

At the very top of the page, you will see a set of tabs:

* **Load / save**: Here you can load or save the model data to a spreadsheet
* **Specification**: Here you can configure the overall model structure
* **Data**: Here you can view fill in the data tables required by the structure
* **Gadget3 script**: Here you can get an R script to build a [gadget3](https://gadget-framework.github.io/gadget3/) model using your specification & data
* **SS3/r4ss script**: Here you can get an R script to build a [r4ss](https://github.com/r4ss/r4ss) model using your specification & data

## Starting a model

By default, Modelwizard will start with all fields empty, so you can start a new model from scratch.

If you want to start from the demo model, click the **Load / save** tab and **Load demo data**

If you want to start from a previously saved model, click the **Load / save** tab and upload a file under **Load spreadsheet**

## Filling in a model specification

Firstly, you need to describe your scenario on the **Specification** tab.
Either start here to fill in details or click "Load demo data" on the **Load / save** tab.

### **Area** section

Modelwizard currently only supports single-area models.
Enter a name for the area your model covers here.

### **Time** section

Enter the years your model will cover, and seasonality if desired.

### **Stock** section

Give your stock a name and describe it's physical attributes, to configure the model bins.

### **Commercial fleets** section

For each commercial fleet present in your model, click **Add new** to add a section to the page:

* **identifier**: A short name for the fleet to use in the model
* **Start/end year for fleet**: If the fleet is not present for the entire model period, enter the years for which it is operational
* **Landings in**: The unit the landings data will be provided in the next tab
* **Length distribution** / **Age-Length distribution**: If data is available for either, choose the unit and provide the data on the next tab

### **Survey fleets** section

For each survey fleet present in your model, click **Add new** to add a section to the page.
The fields are the same as above, apart from:

* **Catch Per Unit Effort (CPUE)**: Unlike commercial fleets, survey fleets are assumed to catch with constant effort. Select the unit the CPUE data will be provided in the next tab

## Providing observation data for the model

The **Data** tab provides a spreadsheet view for each data set required by your model.
Year/step/area/age/length columns will be pre-filled based on the specification, and a value column will be ready to fill in.

If you prefer, you can use Excel to fill in your data. Go back to the **Load / save** tab, click **Save data to spreadsheet**, and open the downloaded spreadsheet in Excel.

Finally, you can skip this step entirely. Instead, generate a model script, and modify the code to load the data from another source.

## Building a gadget3 model

Once all model data is provided, you can visit the **Gadget3 script** tab to download a script that will build and run your model.

Download the spreadsheet, copy-paste the code into a R script and run it.

Visit [demo_output_figures](demo_output_figures.html) to see representative output from running the demo.

At this point you have should have a functional gadget3 model.
However, the results it produces will not be useful without futher tuning.
The next step should be the [gadget3 documentation](https://gadget-framework.github.io/gadget3/articles/) to learn more about how to customise your model.

## Building a SS3 model

Once all model data is provided, you can visit the **SS3/r4ss script** tab to download a script that will build and run your model.

Download the spreadsheet, copy-paste the code into a R script and run it.

At this point you have should have a functional SS3 model.
However, the results it produces will not be useful without futher tuning.
The next step should be the [SS3 documentation](https://vlab.noaa.gov/web/stock-synthesis) and [r4ss documentation](https://r4ss.github.io/r4ss/) to learn more about how to customise your model.

Visit [ss3/plots/_SS_output.html](ss3/plots/_SS_output.html) to see representative output from running the demo.

Also, visit [the Math4Fish website](https://math4fish.ieo.csic.es/proyecto-2-3/) for introductory material on developing an SS3 model.
