---
title: "Math4Fish Entregable TODO"

author:
- Jamie Lentin^[lentinj@shuttlethread.com]
- Alfonso Perez^[alfonso.perez@ieo.csic.es]
- María José Zúñiga^[maria.zuniga@ieo.csic.es]
- María Angeles Gamaza
- Andrea Jiménez
- Margarita María Rincón
output:
  # word_document:
  #   toc: no
  #   toc_depth: '6'
  pdf_document:
    keep_tex: yes
    toc: no
    toc_depth: '4'
    number_sections: yes
header-includes:
- \usepackage{fancyhdr}
- \usepackage{array}
- \usepackage{booktabs}
- \usepackage{threeparttablex}
- \pagestyle{fancyplain}
- \usepackage{caption}
- \usepackage{listings}% http://ctan.org/pkg/listings
- \captionsetup[table]{name=Tabla,labelsep=period}
- \captionsetup[figure]{name=Figura,labelsep=period}
- \captionsetup{justification=justified,format=plain,font=small,margin=1pt}
- \fancyhead{}
- \addtolength{\headheight}{1.5cm}
- \rhead{\includegraphics[width=3.5cm]{input/LOGO_2SN_recortado.png}}
- \usepackage{float}
- \fancyfoot[R]{\thepage}
- \fancyfoot[C]{\includegraphics[height=3cm]{input/Banners_Convenio.jpeg}}
- \fancyfoot[L]{}
 #\chead{\includegraphics[height=2.0cm]{logos_imagenes/Header_logos.jpg}}
- \pagenumbering{gobble}
- \pagenumbering{arabic}
- \usepackage{fvextra}
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
always_allow_html: true
lang: es-ES
geometry: "left=3cm,right=3cm,top=2cm,bottom=4.5cm"
urlcolor: blue
csl: ices-journal-of-marine-science.csl
bibliography: referencias.bib
link-citations: yes
# editor_options:
#   markdown:
#     wrap: 72
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=T, warning=FALSE, message=FALSE,comment="",collapse=TRUE,fig.align="center",fig.pos="h!",eval=F,tidy=TRUE, tidy.opts=list(arrow=TRUE, indent=2))
```


```{r results='hide',include=FALSE,eval=T}
## Librerías requeridas
 paquetes <- c("stringr", "tidyverse", "kableExtra","ggplot2","ggthemes",
               "patchwork","dplyr","reshape","here","r4ss","zoo")
 lapply(paquetes, require, character.only = TRUE)
```


<!-- \mbox{} con esto inserto una línea en blanco -->

\mbox{}

\mbox{}

\mbox{}

Instituto Español de Oceanografía, Consejo Superior de Investigaciones
Científicas (IEO-CSIC)

```{=tex}
\renewcommand{\tablename}{Tabla}
\newpage
```
\mbox{} \setcounter{tocdepth}{5} \tableofcontents \newpage

# TODO

* Deliverable number, name
* Should we call it ModelWizard, or IEOInputSA?

# Introduction

Evaluating several modelling frameworks to decide on the most appropriate for a scenario is a common activity.
This will most likely be undertaken when unfamiliar with modelling frameworks,
and a learning curve will have to be scaled before being able to produce a working model in each, which takes time.

Automating the process, creating equivalent models in different modelling frameworks isn't possible.
There would not be a sufficient overlap in featuresets for this to make sense.
However, a good portion of this work is formatting data into a shape suitable for ingestion by the modelling framework.
In comparison, this can be laborous rote work.

ModelWizard aims to assist with both problems.
After filling in a simple web form and copying data in place, scripts will be generated that can produce Gadget3 and SS3 models.
These aren't complete models in themselves, more customisation will be required.
However, instead of a generic example you can now start with a script that works with your scenario and data.

## Architecture

ModelWizard is built in R using the Shiny framework, this allows rapid development of web applications in R
The Shiny framework is *reactive*, automatically propogating changes from the web browser to the server,
By building in R we can also utilise existing packages in R, such as r4ss and gadget3.

A ModelWizard configuration is loaded/saved to spreadsheet files.
This means no data is stored on the server, in turn requiring access restrictions.
It also means the model data can be opened in Excel, and a more familiar environment used to edit observation data.

ModelWizard doesn't itself run any models, instead R scripts are formed that you can then use on your computer.
Whilst running models would be possible, it would increase server load dramatically.
It also wouldn't be useful, as the idea behind ModelWizard is the produced models are a starting point for experimentation, and not a finished end product.

## Installation

Nothing needs to be installed, ModelWizard is freely available on the internet at http://math4fish.shinyapps.io/modelwizard,
using shinyapps.io for hosting shiny applications.

ModelWizard is open source, and the source code is available at https://github.com/gadget-framework/modelwizard/.
From here you can either install it on your local computer, or install a modified version on shinyapps.io.

# Financiación

Proyecto financiado por la Unión Europea-NextGenerationEU. Componente 3. Inversión 7. Convenio entre el Ministerio de Agricultura, Pesca, y Alimentación y la Agencia Estatal Consejo Superior de Investigaciones Científicas M.P. -a través del Instituto Español de Oceanografía- para impulsar la investigación pesquera como base para la gestión pesquera sostenible.  Eje6, Math4Fish: Nuevas herramientas para el modelado matemático en el asesoramiento científico de pesquerías españolas.

\newpage

# Bibliografía

 - Chang W, Cheng J, Allaire J, Sievert C, Schloerke B, Xie Y, Allen J, McPherson J, Dipert A, Borges B (2024). shiny: Web Application Framework for R. R package version 1.8.1.1, https://CRAN.R-project.org/package=shiny
 - Lentin J, Elvarsson B, Butler W (2024). gadget3: Globally-Applicable Area Disaggregated General Ecosystem Toolbox V3. https://gadget-framework.github.io/gadget3/.
 - Taylor, I.G., Doering, K.L., Johnson, K.F., Wetzel, C.R., Stewart, I.J., 2021. Beyond visualizing catch-at-age models: Lessons learned from the r4ss package about software to support stock assessments. Fisheries Research, 239:105924 https://doi.org/10.1016/j.fishres.2021.105924

<div id="refs"></div>
